Appendix to cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-06-16 jss/appendix.Rmd
appendix with script to create historical_tmax

---
title: "appendix to cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
author: "Sherry Zhang"
date: '2022-06-14'
output: pdf_document
---

@@

# Appendix

@@

\newpage

>>>>

---
title: "Appendix to cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
author: "Sherry Zhang"
date: '2022-06-14'
output: 
  pdf_document:
    number_section: true
---

@@

# Linking pipeline

@@

# Script to obtain example data

The script below presents the codes required to obtain the data, `historical_tmax`, used in the example of Section 5.2 *Australian historical maximum temperature*. The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time. A copy of the data is provided in the data folder of the paper repository at: https://github.com/huizezhang-sherry/paper-cubble.

```{r eval = FALSE}
library(tidyverse)
library(cubble)
all_stations <- rnoaa::ghcnd_stations() |>
  filter(str_starts(id, "ASN")) |> # Australian stations start wiht "ASN"
  filter(last_year >= 2020) |>
  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) |>
  select(-state, -gsn_flag) |>
  select(id, longitude, latitude, elevation, name,
         wmo_id, element, first_year, last_year) |>
  rename(long = longitude, lat = latitude, elev = elevation) |> 
  filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))

raw_tmax <- all_stations |>
  rowwise() |>
  mutate(ts = list(rnoaa::meteo_pull_monitors(
    monitors = id, var = "TMAX",
    date_min = glue::glue("{first_year}-01-01"),
    date_max = glue::glue("{last_year}-12-31")
    ) |>
      select(-id)
    )
  )

historical_tmax <- raw_tmax |>
  select(-element) |>
  unnest(ts) |>
  mutate(tmax = tmax/10) |>
  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) |> 
  as_cubble(index = date, key = id, coords = c(long, lat))

save(historical_tmax, file = here::here("data/historical_tmax.rda"))
```
